### 2019-09-24 manuscript.tex
updates to project from summer

In this chapter, I will focus on the compilation of paleomagnetic poles developed through the Nordic Paleomagnetism Workshops with some additions and modifications. The Nordic Paleomagnetism Workshops have taken the approach of using expert panels to assess paleomagnetic poles and assign them grades meant to convey the confidence that the community has in these results. While many factors associated with paleomagnetic poles can be assessed quantitatively such as the Fisher statistics and the precision of geochronological constraints, other aspects such as the degree to which available field tests constrain the magnetization to be primary require expert assessment. The categorizations used by the expert panel are `A' and `B' with the last panel meeting occuring in Fall 2017 in Leirubakki, Iceland. An `A' rating refers to poles that are judged to be of such high-quality that they provide essential constraints that should be satisfied in paleogeographic reconstructions. A `B' rating is associated with poles that are judged to likely provide a high-quality constraint, but have some deficiency such as remaining ambiguity in the demonstration of primary remanence or the quality/precision of available geochronologic constraints. Additional poles that were not given an `A' or `B' classification at the Nordic Workshops are referred to as not-rated (`NR'). These additional poles are taken from the Paleomagia database \citep{Veikkolainen2013a}

@@

Following Laurentia's amalgamation, poles from each part of the Laurentia can be considered to reflect the position of the entire composite craton. It is worth considering the possibility that poles from zones of Paleoproterozoic and Mesoproterozoic accretion could be allochthonous to the craton. \cite{Hall201Xa} argued that this was the case for late Mesoproterozoic and early Neoproterozoic poles from east of the Grenvillian allochothon boundary fault. However, the majority of researchers have considered these poles to post-date major differential motion and be associated with cooling during collapse of thick plateau developed during continent-continent collision (e.g. \cite{}). Poles with a B-rating are also included in the composite that come from Greenland, Svalbard and Scotland. These terranes were once part of contiguous Laurentia, but have subsequently rifted away. These poles need to be rotated into the Laurentia reference frame prior to use for tectonic reconstruction and I apply the following rotations:

@@

>>>>

In this chapter, I will focus on the compilation of paleomagnetic poles developed through the Nordic Paleomagnetism Workshops with some additions and modifications. The Nordic Paleomagnetism Workshops have taken the approach of using expert panels to assess paleomagnetic poles and assign them grades meant to convey the confidence that the community has in these results. While many factors associated with paleomagnetic poles can be assessed quantitatively such as the Fisher statistics and the precision of geochronological constraints, other aspects such as the degree to which available field tests constrain the magnetization to be primary require expert assessment. The categorizations used by the expert panel are `A' and `B' with the last panel meeting occurring in Fall 2017 in Leirubakki, Iceland. An `A' rating refers to poles that are judged to be of such high-quality that they provide essential constraints that should be satisfied in paleogeographic reconstructions. A `B' rating is associated with poles that are judged to likely provide a high-quality constraint, but have some deficiency such as remaining ambiguity in the demonstration of primary remanence or the quality/precision of available geochronologic constraints. Additional poles that were not given an `A' or `B' classification at the Nordic Workshops are referred to as not-rated (`NR'). These additional poles are taken from the Paleomagia database \citep{Veikkolainen2013a}

@@

Following Laurentia's amalgamation, poles from each part of the Laurentia can be considered to reflect the position of the entire composite craton. It is worth considering the possibility that poles from zones of Paleoproterozoic and Mesoproterozoic accretion could be allochthonous to the craton. \cite{Halls201Xa} argued that this was the case for late Mesoproterozoic and early Neoproterozoic poles from east of the Grenvillian allochothon boundary fault. However, the majority of researchers have considered these poles to post-date major differential motion and be associated with cooling during collapse of a thick orogenic plateau developed during continent-continent collision (e.g. \citealp{Brown2012a}}). Poles with a B-rating are also included in the composite that come from Greenland, Svalbard and Scotland. These terranes were once part of contiguous Laurentia, but have subsequently rifted away. These poles need to be rotated into the Laurentia reference frame prior to use for tectonic reconstruction and I apply the following rotations:

@@

The synthesis of \citet{Pesonen2012a} developed reconstructions at various Proterozoic time slices.
